Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-02-26 index.qmd
switch to 2021 CLB data

```{python}
from lymph import models

graph_dict = {
  ("tumor", "T"): ["II", "III"],
  ("lnl", "II"): ["III"],
  ("lnl", "III"): [],
}
model = models.Midline(
  graph_dict=graph_dict,
  is_symmetric={"lnl_spread": True},
  use_mixing=True,
  use_central=False,
  use_midext_evo=True,
)
model.set_modality("max_llh", spec=1.0, sens=1.0)
```

>>>>

# Methods

Prepare the model and load the data:

```{python}
import numpy as np
import pandas as pd
from emcee import EnsembleSampler, backends, moves
from corner import corner

from lymph import models
from lyscripts.utils import binom_pmf, parametric_binom_pmf

np.random.seed(42)
```

```{python}
graph_dict = {
  ("tumor", "T"): ["II", "III"],
  ("lnl", "II"): ["III"],
  ("lnl", "III"): [],
}
model = models.Midline(
  graph_dict=graph_dict,
  is_symmetric={"lnl_spread": True},
  use_mixing=True,
  use_central=False,
  use_midext_evo=True,
)
model.set_modality("max_llh", spec=1.0, sens=1.0)
model.set_distribution(
  "early",
  binom_pmf(np.arange(model.max_time+1), model.max_time, p=0.3),
)
model.set_distribution("late", parametric_binom_pmf)
model.load_patient_data(
  pd.read_csv("data/enhanced.csv", header=[0,1,2]),
)
```

Do the sampling:

```{python}
ndim = model.get_num_dims()
initial = np.random.uniform(size=(20 * ndim, ndim))

backend = backends.HDFBackend(filename="models/midline.hdf5")
moves_mix = [(moves.DEMove(), 0.8), (moves.DESnookerMove(), 0.2)]
sampler = EnsembleSampler(
  nwalkers=20 * ndim,
  ndim=ndim,
  log_prob_fn=model.likelihood,
  backend=backend,
  moves=moves_mix,
  parameter_names=list(model.get_params().keys())[:ndim],
)
sampler.run_mcmc(initial, 1000, progress=True)
```

Look at the results:

```{python}
backend = backends.HDFBackend(
  filename="models/midline.hdf5",
  read_only=True,
)

samples = backend.get_chain(discard=500, thin=10, flat=True)
corner(
  samples,
  titles=list(model.get_params().keys()),
  show_titles=True,
);
```

```{python}
mean_samples = {
  k: s.mean() for k, s in zip(model.get_params(), samples.T)
}
model.set_params(**mean_samples)
```

NOTE: For the `Bilateral` model the results look absolutely plausible, but the `Midline` model seems to have some issues (`late_p` was learned to be 0.17 !?).
